---
title: "Covid91 vaccine study Final2023"
author: "Yessica Garay!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


$H_0$ There is no difference in the effectiness of COVID921 between females..

$H_0$ There is a difference in the effectiness of COVID921 between female.

## Methods
variable we are going to use

Males: Categorical
Females: Categorical
LGBT: Categorical
Druggies: Categorical

Efecctiness of the vaccine depend on the treatment among the
Subgroups of population.


##Descriptive Result

### Graphical Descriptive Results

##Females

```{r}
dataF = filter(FinalData,sex=="F")


barchartGC(~infected + treatment,data=dataF)
barchartGC(~infected + treatment,data=dataF, type= "percent")

```

Based on the chart there is a difference between the placebo and the covid 19 vaccine.

##Numerical

```{r}
table1 <- xtabs(~infected + treatment, data=dataF)
rowPerc(table1)
colPerc(table1)
```


From this data, it seems that the actual vaccine has a lower infection rate compared to the placebo vaccine, and among those infected, the actual vaccine has a slightly better treatment outcome.


##Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


Fisher's Exact Test suggests that the odds of being infected with COVID-19 are approximately 1.49 times higher with the placebo compared to the vaccine. There is a big association between the Covid19 and vaccine.


##Conclusion

those chart provides strong evidence that the vaccine is effective in reducing the risk of COVID-19 infection compared to the placebo.so we reject 







## LGBTQ



## Druggies


# Overall Results and Conclusions