---
title: "Covid91 vaccine study Final2023"
author: "Yessica Garay!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


$H_0$ There is no difference in the effectiness of COVID921 between females..

$H_0$ There is a difference in the effectiness of COVID921 between female.

## Methods
variable we are going to use

Males: Categorical
Females: Categorical
LGBT: Categorical
Druggies: Categorical

Efecctiness of the vaccine depend on the treatment among the
Subgroups of population.


##Descriptive Result

### Graphical Descriptive Results

##Females

```{r}
dataF = filter(FinalData,sex=="F")


barchartGC(~infected + treatment,data=dataF)
barchartGC(~infected + treatment,data=dataF, type= "percent")

```

Based on the chart there is a difference between the placebo and the covid 19 vaccine.

##Numerical

```{r}
table1 <- xtabs(~infected + treatment, data=dataF)
rowPerc(table1)
colPerc(table1)
```


From this data, it seems that the actual vaccine has a lower infection rate compared to the placebo vaccine, and among those infected, the actual vaccine has a slightly better treatment outcome.


## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
```


Fisher's Exact Test suggests that the odds of being infected with COVID-19 are approximately 1.49 times higher with the placebo compared to the vaccine. There is a big association between the Covid19 and vaccine.


##Conclusion

those chart provides strong evidence that the vaccine is effective in reducing the risk of COVID-19 infection compared to the placebo.so we reject 


## LGBTQ


##Intrdotuction

$H_0$ There is no difference in the effectiness of COVID921 between Gay.

$H_0$ There is a difference in the effectiness of COVID921 between Gay.

##Methods

LGBT/Gays Categorical

##Descriptive Result

### Graphical Descriptive Results

```{r}
levels(FinalData$LGBTQ)
dataG = filter(FinalData,LGBTQ=="gay")
            


barchartGC(~infected + treatment,data=dataG)
barchartGC(~infected + treatment,data=dataG, type= "percent")
```

This chart shows there is significant different between the placebo end vaccine.
among those that got infected the chart show the hight percentage got infected with covid19 with the vaccine.

##Numerical

```{r}
table1 <- xtabs(~infected + treatment, data=dataG)
rowPerc(table1)
colPerc(table1)
```

The infection rate among those who received the placebo is 15.38%, while among those who received the vaccine, it's only 0.66%. This suggests that the vaccine is effective in reducing the risk of COVID-19 infection.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
```
##Conclusion


With such a low p-value, 0 we reject the null hypothesis and conclude that there is a significant association between treatment type and infection status. In other words, it suggests that the type of treatment (placebo or vaccine) is related to of getting infected with COVID-19.


## Druggies


##Intrdotuction

$H_0$ There is no difference in the effectiness of COVID921 Druggies.

$H_0$ There is a difference in the effectiness of COVID921 Druggies.


##Methods

Druggies/Drug use categorical

```{r}
levels(FinalData$DrugUser)
dataD = filter(FinalData,DrugUser=="yes")
            


barchartGC(~infected + treatment,data=dataD) 
barchartGC(~infected + treatment,data=dataD, type= "percent")
```

The chart *ow us the is a significant difference between the druguser " Yes "   among those the are OK 


##Numerical

```{r}
table1 <- xtabs(~infected + treatment, data=dataD)
rowPerc(table1)
colPerc(table1)
```

COVID-19 cases occurred among individuals who received the vaccine, a higher percentage of vaccine recipients remained uninfected compared to the placebo group.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
```


Similar to the previous chart a low p-value, we reject the null hypothesis and conclude that there is a significant association between treatment type and infection status in this dataset as well. placebo and vaccine got infected with covid19




# Overall Results and Conclusions

 it can be concluded that there is a significant association between the type of treatment (placebo vs. vaccine) and infection status (COVID-19 infected vs. not infected) in the studied groups.Therefore, it's reasonable to conclude that the type of treatment (placebo or vaccine) has a substantial impact on the likelihood of being infected with COVID-19 in the studied groups.

Therefore we reject the null hypothessis there significate differences between the groups of females, LGBT/Gays and DrugUser/yes. and conclude that there is a significant association between treatment type and infection status in the studied groups.





 